Visual Diagnostics for Constrained Optimisation with Application to Guided Tours

### 2021-03-08 paper/zhang-cook-laa-langrene-menendez.Rmd
small changes flagged by Gael in the confirmation

---
title: "Visual Diagnostics for Constrained Optimisation with Application to Guided Tours"
author:
  - name: H.Sherry Zhang
    affiliation: Monash University 
    address: Department of Econometrics and Business Statistics
    email:  huize.zhang@monash.edu
  - name: Dianne Cook
    affiliation: Monash University 
    address: Department of Econometrics and Business Statistics
    email:  dicook@monash.edu
  - name: Ursula Laa
    affiliation: University of Natural Resources and Life Sciences
    address: Institute of Statistics
    email:  ursula.laa@boku.ac.at  
  - name: Nicolas Langrené
    affiliation: CSIRO Data61 
    address: 34 Village Street, Docklands VIC 3008 Australia
    email: nicolas.langrene@csiro.au
  - name: Patricia Menéndez
    affiliation: Monash University 
    address: Department of Econometrics and Business Statistics
    email:  patricia.menendez@monash.edu 
abstract: |
  Guided tour searches for interesting low-dimensional views of high-dimensional data via optimising a projection pursuit index function. The first paper of projection pursuit by @friedman1974projection stated that "the technique used for maximising the projection index strongly influences both the statistical and the computational aspects of the procedure." While much work has been done in proposing  indices in the literature, less has been done on evaluating the performance of the optimisers. In this paper, we implement a data collection object in the optimisation of projection pursuit guided tour and introduce visual diagnostics based on the data object collected. These diagnostics and workflow can be applied to a broad class of optimisers, to assess their performance. An R package, \pkg{ferrn}, has been created to implement the diagnostics.

@@

Optimization problems are ubiquitous in many areas of study. While in some cases analytical solutions can be found, the majority of problems rely on numerical methods to find the optimal solution. These numerical methods follow iterative approaches that aim at finding the optimum by progressively improving the current solution until a desirable accuracy is achieved. Although this principle seems uncomplicated, a number of challenges arise such as the possible existence of multiple maxima (local and global), constraints and noisy objective function, and the trade-off between desirable accuracy and computational burden. In addition, the optimization results might depend on the algorithm starting values, affecting the consistency of results.

@@

```{r toy-interp, fig.cap = "An inspection of the index values as the optimisation progress for two optimisers: CRS (left) and SA (right). The holes index is optimised for a 2D projection problem on the six-variable dataset \\code{boa6}. Lines indicate the interpolation and dots indicate new target bases generated by the optimisers. Interpolation in both optimisation is smooth while SA is observed to first pass by some bases with higher index value before reach the target bases in Time 76-130."}
```

@@

```{r polish, fig.height= 5, fig.cap = "Comparison of the projected data before and after using polishing along with the trace of index value of a 2D projection problem on \\code{boa6} data using holes index. Optimiser CRS is used before the polish. The clustering structure in the data is detected by CRS but the polish step improves the index value and produces clearer boundaries of the clusters, espeically along the vertial axis."}
```

>>>>

---
title: "Visual Diagnostics for Constrained Optimisation with Application to Guided Tours"
author:
  - name: H.Sherry Zhang
    affiliation: Monash University 
    address: Department of Econometrics and Business Statistics
    email:  huize.zhang@monash.edu
  - name: Dianne Cook
    affiliation: Monash University 
    address: Department of Econometrics and Business Statistics
    email:  dicook@monash.edu
  - name: Ursula Laa
    affiliation: University of Natural Resources and Life Sciences
    address: Institute of Statistics
    email:  ursula.laa@boku.ac.at  
  - name: Nicolas Langrené
    affiliation: CSIRO Data61 
    address: 34 Village Street, Docklands VIC 3008 Australia
    email: nicolas.langrene@csiro.au
  - name: Patricia Menéndez
    affiliation: Monash University 
    address: Department of Econometrics and Business Statistics
    email:  patricia.menendez@monash.edu 
abstract: |
  Guided tour searches for interesting low-dimensional views of high-dimensional data via optimising a projection pursuit index function. The first paper of projection pursuit by @friedman1974projection stated that "the technique used for maximising the projection index strongly influences both the statistical and the computational aspects of the procedure." While much work has been done in proposing  indices in the literature, less has been done on evaluating the performance of the optimisers. In this paper, we implement a data collection object in the optimisation of projection pursuit guided tour and introduce visual diagnostics based on the data object collected. These diagnostics and this workflow can be applied to a broad class of optimisers, to assess their performance. An R package, \pkg{ferrn}, has been created to implement the diagnostics.

@@

Optimisation problems are ubiquitous in many areas of study. While in some cases analytical solutions can be found, the majority of problems rely on numerical methods to find the optimal solution. These numerical methods follow iterative approaches that aim at finding the optimum by progressively improving the current solution until a desirable accuracy is achieved. Although this principle seems uncomplicated, a number of challenges arise such as the possible existence of multiple maxima (local and global), constraints and noisy objective function, and the trade-off between desirable accuracy and computational burden. In addition, the optimization results might depend on the algorithm starting values, affecting the consistency of results.

@@

```{r toy-interp, fig.cap = "An inspection of the index values as the optimisation progress for two optimisers: CRS (left) and SA (right). The holes index is optimised for a 2D projection problem on the six-variable dataset \\code{boa6}. Lines indicate the interpolation and dots indicate new target bases generated by the optimisers. Interpolation in both optimisation is smooth while SA is observed to first pass by some bases with higher index value before reach the target bases in time 76-130."}
```

@@

```{r polish, fig.height= 5, fig.cap = "Comparison of the projected data before and after using polishing along with the trace of index value of a 2D projection problem on \\code{boa6} data using holes index. Optimiser CRS is used before the polish. The clustering structure in the data is detected by CRS but the polish step improves the index value and produces clearer boundaries of the clusters, especially along the vertial axis."}
```

### 2021-03-08 paper/zhang-cook-laa-langrene-menendez.tex
small changes flagged by Gael in the confirmation

\abstract{%
Guided tour searches for interesting low-dimensional views of
high-dimensional data via optimising a projection pursuit index
function. The first paper of projection pursuit by
\citet{friedman1974projection} stated that ``the technique used for
maximising the projection index strongly influences both the statistical
and the computational aspects of the procedure.'' While much work has
been done in proposing indices in the literature, less has been done on
evaluating the performance of the optimisers. In this paper, we
implement a data collection object in the optimisation of projection
pursuit guided tour and introduce visual diagnostics based on the data
object collected. These diagnostics and workflow can be applied to a
broad class of optimisers, to assess their performance. An R package,
\pkg{ferrn}, has been created to implement the diagnostics.

@@

\caption[An inspection of the index values as the optimisation progress for two optimisers]{An inspection of the index values as the optimisation progress for two optimisers: CRS (left) and SA (right). The holes index is optimised for a 2D projection problem on the six-variable dataset \code{boa6}. Lines indicate the interpolation and dots indicate new target bases generated by the optimisers. Interpolation in both optimisation is smooth while SA is observed to first pass by some bases with higher index value before reach the target bases in Time 76-130.}\label{fig:toy-interp}
\end{figure}
\end{Schunk}

@@

\caption[Comparison of the projected data before and after using polishing along with the trace of index value of a 2D projection problem on \code{boa6} data using holes index]{Comparison of the projected data before and after using polishing along with the trace of index value of a 2D projection problem on \code{boa6} data using holes index. Optimiser CRS is used before the polish. The clustering structure in the data is detected by CRS but the polish step improves the index value and produces clearer boundaries of the clusters, espeically along the vertial axis.}\label{fig:polish}
\end{figure}
\end{Schunk}

>>>>

\abstract{%
Guided tour searches for interesting low-dimensional views of
high-dimensional data via optimising a projection pursuit index
function. The first paper of projection pursuit by
\citet{friedman1974projection} stated that ``the technique used for
maximising the projection index strongly influences both the statistical
and the computational aspects of the procedure.'' While much work has
been done in proposing indices in the literature, less has been done on
evaluating the performance of the optimisers. In this paper, we
implement a data collection object in the optimisation of projection
pursuit guided tour and introduce visual diagnostics based on the data
object collected. These diagnostics and this workflow can be applied to
a broad class of optimisers, to assess their performance. An R package,
\pkg{ferrn}, has been created to implement the diagnostics.

@@

\caption[An inspection of the index values as the optimisation progress for two optimisers]{An inspection of the index values as the optimisation progress for two optimisers: CRS (left) and SA (right). The holes index is optimised for a 2D projection problem on the six-variable dataset \code{boa6}. Lines indicate the interpolation and dots indicate new target bases generated by the optimisers. Interpolation in both optimisation is smooth while SA is observed to first pass by some bases with higher index value before reach the target bases in time 76-130.}\label{fig:toy-interp}
\end{figure}
\end{Schunk}

@@

\caption[Comparison of the projected data before and after using polishing along with the trace of index value of a 2D projection problem on \code{boa6} data using holes index]{Comparison of the projected data before and after using polishing along with the trace of index value of a 2D projection problem on \code{boa6} data using holes index. Optimiser CRS is used before the polish. The clustering structure in the data is detected by CRS but the polish step improves the index value and produces clearer boundaries of the clusters, especially along the vertial axis.}\label{fig:polish}
\end{figure}
\end{Schunk}
